---
title: "Netflix Content Analysis"
author: "Veni_Neha_Minha_Niyukta"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: darkly
    toc: true
    toc_float: true
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE,
                      fig.width=9, fig.height=5, fig.align="center")
```

---

## 1. Load Packages
```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
pkgs <- c("ggplot2","dplyr","tidyr","stringr","scales")
for(p in pkgs){
  if(!requireNamespace(p, quietly=TRUE)) install.packages(p)
  library(p, character.only=TRUE)
}
```

---

## 2. Load Data
```{r}
titles  <- read.csv("titles.csv",  stringsAsFactors=FALSE, na.strings=c("","NA"))
credits <- read.csv("credits.csv", stringsAsFactors=FALSE, na.strings=c("","NA"))

cat("Titles:", nrow(titles), "\n")
cat("Credits:", nrow(credits), "\n")
head(titles[, c("title","type","release_year","genres","imdb_score","age_certification")], 5)
```

---

## 3. Preprocessing
```{r}
# Check missing values
cat("Missing values:\n")
colSums(is.na(titles[, c("imdb_score","age_certification","genres","production_countries")]))

# Fill missing values
titles$imdb_score[is.na(titles$imdb_score)] <- median(titles$imdb_score, na.rm=TRUE)
titles$age_certification[is.na(titles$age_certification)] <- "Not Rated"

# Clean genres and countries columns
titles$genres <- gsub("\\[|\\]|\\'", "", titles$genres)
titles$production_countries <- gsub("\\[|\\]|\\'", "", titles$production_countries)

# Keep only MOVIE and SHOW
titles <- titles %>% filter(type %in% c("MOVIE","SHOW"), !is.na(title))

cat("\nAfter cleaning:", nrow(titles), "titles\n")
cat("Missing imdb_score now:", sum(is.na(titles$imdb_score)), "\n")
```

---

## 4. Summary Statistics
```{r}
summary(titles[, c("imdb_score","release_year")])

titles %>%
  group_by(type) %>%
  summarise(
    Count       = n(),
    Mean_IMDB   = round(mean(imdb_score, na.rm=TRUE), 2),
    Median_IMDB = round(median(imdb_score, na.rm=TRUE), 2),
    SD_IMDB     = round(sd(imdb_score, na.rm=TRUE), 2),
    Min_IMDB    = min(imdb_score, na.rm=TRUE),
    Max_IMDB    = max(imdb_score, na.rm=TRUE)
  ) %>%
  knitr::kable(caption="Summary Statistics by Content Type")
```

**Conclusion:**
Movies and TV Shows differ in their average IMDB scores. The table above
shows the spread and central tendency of ratings across both content types.

---

## 5. Movies vs Shows
```{r}
type_count <- titles %>%
  count(type) %>%
  mutate(pct   = n/sum(n),
         label = paste0(type,"\n",comma(n),"\n(",percent(pct,accuracy=1),")"))

ggplot(type_count, aes(x="", y=n, fill=type)) +
  geom_col(width=1, color="white") +
  coord_polar(theta="y") +
  geom_text(aes(label=label), position=position_stack(vjust=0.5),
            size=5, fontface="bold", color="white") +
  scale_fill_manual(values=c("MOVIE"="#E50914","SHOW"="#221F1F")) +
  labs(title="Movies vs TV Shows on Netflix") +
  theme_void() +
  theme(plot.title=element_text(face="bold",size=16,hjust=0.5,color="#E50914"),
        legend.position="none")
```

**Conclusion:**
Netflix has significantly more Movies than TV Shows. This shows that Netflix
built its library primarily around films before expanding into original series.

---

## 6. Top Genres
```{r}
genres_df <- titles %>%
  filter(!is.na(genres), genres!="") %>%
  separate_rows(genres, sep=",") %>%
  mutate(genres=str_trim(genres)) %>%
  filter(genres!="") %>%
  count(genres, sort=TRUE) %>%
  slice_head(n=15)

ggplot(genres_df, aes(x=reorder(genres,n), y=n, fill=n)) +
  geom_col(show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.15, size=3.5, fontface="bold") +
  coord_flip() +
  scale_fill_gradient(low="#FF6B6B", high="#E50914") +
  scale_y_continuous(expand=expansion(mult=c(0,0.15))) +
  labs(title="Top 15 Genres on Netflix", x=NULL, y="Number of Titles") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        panel.grid.major.y=element_blank())
```

**Conclusion:**
Drama and Comedy are the most dominant genres on Netflix. This reflects
global audience preference for emotionally engaging and light-hearted content.

---

## 7. Content Added Per Year
```{r}
year_df <- titles %>%
  filter(!is.na(release_year), release_year>=2000, release_year<=2023) %>%
  count(release_year, type)

ggplot(year_df, aes(x=release_year, y=n, color=type, group=type)) +
  geom_line(linewidth=1.2) +
  geom_point(size=2) +
  scale_color_manual(values=c("MOVIE"="#E50914","SHOW"="#F5A623")) +
  scale_x_continuous(breaks=seq(2000,2023,2)) +
  labs(title="Content Released Per Year (2000-2023)",
       x="Year", y="Number of Titles", color="Type") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        legend.position="top")
```

**Conclusion:**
Netflix content production increased sharply after 2015. This reflects
Netflix's global expansion strategy and heavy investment in original content.

---

## 8. Top Countries
```{r}
countries_df <- titles %>%
  filter(!is.na(production_countries), production_countries!="") %>%
  separate_rows(production_countries, sep=",") %>%
  mutate(production_countries=str_trim(production_countries)) %>%
  filter(production_countries!="") %>%
  count(production_countries, sort=TRUE) %>%
  slice_head(n=15)

ggplot(countries_df, aes(x=reorder(production_countries,n), y=n, fill=n)) +
  geom_col(show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.15, size=3.5, fontface="bold") +
  coord_flip() +
  scale_fill_gradient(low="#FFC3C3", high="#E50914") +
  scale_y_continuous(expand=expansion(mult=c(0,0.15))) +
  labs(title="Top 15 Countries Producing Netflix Content", x=NULL, y="Titles") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        panel.grid.major.y=element_blank())
```

**Conclusion:**
The United States produces the most Netflix content by far, followed by
India and the UK. This highlights Netflix's dominance in English-language
and South Asian markets.

---

## 9. IMDB Score Distribution
```{r}
imdb_df <- titles %>% filter(!is.na(imdb_score), imdb_score>0)

ggplot(imdb_df, aes(x=imdb_score, fill=type)) +
  geom_histogram(binwidth=0.5, color="white", alpha=0.85, position="identity") +
  geom_vline(xintercept=median(imdb_df$imdb_score),
             color="yellow", linetype="dashed", linewidth=1) +
  scale_fill_manual(values=c("MOVIE"="#E50914","SHOW"="#F5A623")) +
  scale_x_continuous(breaks=1:10) +
  labs(title="IMDB Score Distribution",
       subtitle="Yellow dashed line = median score",
       x="IMDB Score", y="Count", fill="Type") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        legend.position="top")
```

**Conclusion:**
Most Netflix titles score between 6.0 and 7.0 on IMDB with a median around
6.5. This suggests Netflix hosts mostly average to good quality content
rather than exclusively critically acclaimed titles.

---

## 10. Top Actors & Directors
```{r}
top_actors <- credits %>%
  filter(role=="ACTOR", !is.na(name)) %>%
  count(name, sort=TRUE) %>%
  slice_head(n=15)

ggplot(top_actors, aes(x=reorder(name,n), y=n, fill=n)) +
  geom_col(show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.2, size=3.5, fontface="bold") +
  coord_flip() +
  scale_fill_gradient(low="#FFC3C3", high="#E50914") +
  scale_y_continuous(expand=expansion(mult=c(0,0.15))) +
  labs(title="Top 15 Most Featured Actors", x=NULL, y="Titles") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        panel.grid.major.y=element_blank())
```

**Conclusion:**
A small group of actors appear across many Netflix titles, suggesting
Netflix repeatedly works with familiar and trusted talent.
```{r}
top_directors <- credits %>%
  filter(role=="DIRECTOR", !is.na(name)) %>%
  count(name, sort=TRUE) %>%
  slice_head(n=15)

ggplot(top_directors, aes(x=reorder(name,n), y=n, fill=n)) +
  geom_col(show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.2, size=3.5, fontface="bold") +
  coord_flip() +
  scale_fill_gradient(low="#FFD580", high="#F5A623") +
  scale_y_continuous(expand=expansion(mult=c(0,0.15))) +
  labs(title="Top 15 Most Active Directors", x=NULL, y="Titles") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        panel.grid.major.y=element_blank())
```

**Conclusion:**
A handful of directors are responsible for a large share of Netflix content,
showing Netflix's preference for working with proven and prolific filmmakers.

---

## 11. Age Rating Breakdown
```{r}
rating_df <- titles %>%
  filter(!is.na(age_certification), age_certification!="Not Rated") %>%
  count(age_certification, type)

ggplot(rating_df, aes(x=reorder(age_certification,n), y=n, fill=type)) +
  geom_col(position="dodge") +
  geom_text(aes(label=n), position=position_dodge(width=0.9),
            vjust=-0.4, size=3.5, fontface="bold") +
  scale_fill_manual(values=c("MOVIE"="#E50914","SHOW"="#F5A623")) +
  scale_y_continuous(expand=expansion(mult=c(0,0.15))) +
  labs(title="Content by Age Rating", x="Rating", y="Titles", fill="Type") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"),
        legend.position="top")
```

**Conclusion:**
TV-MA and R ratings dominate Netflix, confirming that the platform primarily
targets adult audiences while still offering family-friendly content.

---

## 12. Regression Model
```{r}
model <- lm(imdb_score ~ release_year, data=titles)
summary(model)

ggplot(titles %>% filter(!is.na(imdb_score), !is.na(release_year)),
       aes(x=release_year, y=imdb_score)) +
  geom_point(alpha=0.2, color="#E50914") +
  geom_smooth(method="lm", color="yellow", linewidth=1.2) +
  labs(title="Regression: IMDB Score vs Release Year",
       x="Release Year", y="IMDB Score") +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold",size=16,color="#E50914"))
```

**Conclusion:**
The regression model shows the relationship between release year and IMDB
score. A negative slope means newer content tends to score slightly lower,
possibly due to the large volume of average content added in recent years.
Release year alone is a weak predictor of quality, meaning other factors
like genre and production budget matter more.

---

*Analysis by Veni_Neha_Minha_Niyukta | Data: Netflix Titles by Victor Soirro (Kaggle)*